---
title: "Heightloss"
author: "Katarina Matthes"
date: "`r Sys.Date()`"
output:
  rmdformats::robobook:
    code_folding: show
    self_contained: true
    highlight: pygments
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

# Table

SITAR_size_velocity not true values!

```{r Tablefull, echo=FALSE,results = "asis",message=FALSE}
tab1 <- tableby(sex ~ height02+height07+height11+height15+SITAR_size_height+SITAR_size_tempo+SITAR_size_velocity+height36+height69+heightdiff+Excess.weight+
                  education43+socialclass+smoking+exercise+fruits+osteo+health.score.sum+chronicDiseaseScore+pain.walking, data=dat,test=FALSE)
summary(tab1)
```

# Association between height in childhood and height loss
Between the ages of 36 and 69, women lost an average of 2.37 cm in height and men 1.81 cm. Women lost significantly more height than men.

two-sample, two-sided, unpaired T-test

```{r}
function_diff_sex()
```

Summary:
<br />
SITAR Height and height at age 36 are significant associated with height loss. Furthermore, none of the co-factors is significantly associated with height loss in men. In women, excess weight is significantly associated with height loss.

## Age 2
Height at age 2 has no effect on height loss (unadjusted and adjusted)

### Male
Unadjusted
```{r}
function_heightgain_loss(varAge="height02",adjusted="no",sex="male")
```

Adjusted
```{r}
function_heightgain_loss(varAge="height02",adjusted="yes",sex="male")
```

### Female
Unadjusted
```{r}
function_heightgain_loss(varAge="height02",adjusted="no",sex="female")
```

Adjusted
```{r}
function_heightgain_loss(varAge="height02",adjusted="yes",sex="female")
```


## Age 7
Height at age 7 has no effect on height loss (unadjusted and adjusted)

### Male
Unadjusted
```{r}
function_heightgain_loss(varAge="height07",adjusted="no",sex="male")
```

Adjusted
```{r}
function_heightgain_loss(varAge="height07",adjusted="yes",sex="male")
```

### Female
Unadjusted
```{r}
function_heightgain_loss(varAge="height07",adjusted="no",sex="female")
```

Adjusted
```{r}
function_heightgain_loss(varAge="height07",adjusted="yes",sex="female")
```


## Delta between age 2 and age 7
Delta height between age 2 and 7 has no effect on height loss (unadjusted and adjusted).

### Male
Unadjusted
```{r}
function_heightgain_loss(varAge="heightgain2_7",adjusted="no",sex="male")
```

Adjusted
```{r}
function_heightgain_loss(varAge="heightgain2_7",adjusted="yes",sex="male")
```

### Female
Unadjusted
```{r}
function_heightgain_loss(varAge="heightgain2_7",adjusted="no",sex="female")
```

Adjusted
```{r}
function_heightgain_loss(varAge="heightgain2_7",adjusted="yes",sex="female")
```

## Age 11
Height at age 11 has no effect on height loss (unadjusted and adjusted)

### Male
Unadjusted
```{r}
function_heightgain_loss(varAge="height11",adjusted="no",sex="male")
```

Adjusted
```{r}
function_heightgain_loss(varAge="height11",adjusted="yes",sex="male")
```

### Female
Unadjusted
```{r}
function_heightgain_loss(varAge="height11",adjusted="no",sex="female")
```

Adjusted
```{r}
function_heightgain_loss(varAge="height11",adjusted="yes",sex="female")
```

## Delta between age 7 and age 11
Delta height between age 7 and 11 has no effect on height loss (unadjusted and adjusted).

### Male
Unadjusted
```{r}
function_heightgain_loss(varAge="heightgain7_11",adjusted="no",sex="male")
```

Adjusted
```{r}
function_heightgain_loss(varAge="heightgain7_11",adjusted="yes",sex="male")
```

### Female
Unadjusted
```{r}
function_heightgain_loss(varAge="heightgain7_11",adjusted="no",sex="female")
```

Adjusted
```{r}
function_heightgain_loss(varAge="heightgain7_11",adjusted="yes",sex="female")
```

## Age 15
Height at age 15 has no effect on height loss (unadjusted and adjusted)

### Male
Unadjusted
```{r}
function_heightgain_loss(varAge="height15",adjusted="no",sex="male")
```

Adjusted
```{r}
function_heightgain_loss(varAge="height15",adjusted="yes",sex="male")
```

### Female
Unadjusted
```{r}
function_heightgain_loss(varAge="height15",adjusted="no",sex="female")
```

Adjusted
```{r}
function_heightgain_loss(varAge="height15",adjusted="yes",sex="female")
```

## SITAR Height
SITAR Height is significantly associated with later height loss. The smaller SITAR height, the less the height loss in adulthood.The result is still significant even after adjustment.

### Male
Unadjusted
```{r}
function_heightgain_loss(varAge="SITAR_size_height",adjusted="no",sex="male")
```

Adjusted
```{r}
function_heightgain_loss(varAge="SITAR_size_height",adjusted="yes",sex="male")
```

### Female
Unadjusted
```{r}
function_heightgain_loss(varAge="SITAR_size_height",adjusted="no",sex="female")
```

Adjusted
```{r}
function_heightgain_loss(varAge="SITAR_size_height",adjusted="yes",sex="female")
```

## SITAR Tempo
SITAR Tempo has no effect on height loss (unadjusted and adjusted).

### Male
Unadjusted
```{r}
function_heightgain_loss(varAge="SITAR_size_tempo",adjusted="no",sex="male")
```

Adjusted
```{r}
function_heightgain_loss(varAge="SITAR_size_tempo",adjusted="yes",sex="male")
```

### Female
Unadjusted
```{r}
function_heightgain_loss(varAge="SITAR_size_tempo",adjusted="no",sex="female")
```

Adjusted
```{r}
function_heightgain_loss(varAge="SITAR_size_tempo",adjusted="yes",sex="female")
```

## SITAR Velocity - NOT REAL DATA FOR NOW
Interpretation after running the model with real data

### Male
Unadjusted
```{r}
function_heightgain_loss(varAge="SITAR_size_velocity",adjusted="no",sex="male")
```

Adjusted
```{r}
function_heightgain_loss(varAge="SITAR_size_velocity",adjusted="yes",sex="male")
```

### Female
Unadjusted
```{r}
function_heightgain_loss(varAge="SITAR_size_velocity",adjusted="no",sex="female")
```

Adjusted
```{r}
function_heightgain_loss(varAge="SITAR_size_velocity",adjusted="yes",sex="female")
```

## Age 36
Height at age 36 is significantly associated with later height loss. The smaller the adult is at age 36, the less the height loss in adulthood. The result is still significant even after adjustment.

### Male
Unadjusted
```{r}
function_heightgain_loss(varAge="height36",adjusted="no",sex="male")
```

Adjusted
```{r}
function_heightgain_loss(varAge="height36",adjusted="yes",sex="male")
```

### Female
Unadjusted
```{r}
function_heightgain_loss(varAge="height36",adjusted="no",sex="female")
```

Adjusted
```{r}
function_heightgain_loss(varAge="height36",adjusted="yes",sex="female")
```

# Association between height loss and health in later life

ACHTUNG: SITAR_size_velocity  is not real! All results of the regression will change after including the true value, but the direction of the results will probably not change!
<!-- <br /> -->
<!-- General question: Leads height loss to health problems or health problems to height loss? In these models height loss leads to health problems. But is this direction true -->
<br />
<br />
Level of education and social class are not significantly related to health outcomes. For women, excess weight and lifestyle factors play a greater role in health outcomes than for men.

## Summary health groups
Men: "Summary health groups" is associated with height loss,  even after adjustment. Smoker and men with excess weight show a significantly higher risk for chronic diseases, than non-smoking men and non-excess weight men. 
<br />
<br />
Women:  "Summary health groups" is associated with height loss,  even after adjustment Women doing exercise show a lower risk of bad general health compared to women doing no exercise.

### Male
Unadjusted
```{r}
function_health_outcome(varOutcome = "health.score.sum",adjusted="no",sex="male")
```

adjusted
```{r}
function_health_outcome(varOutcome = "health.score.sum",adjusted="yes",sex="male")
```

### Female
Unadjusted
```{r}
function_health_outcome(varOutcome = "health.score.sum",adjusted="no",sex="female")
```

adjusted
```{r}
function_health_outcome(varOutcome = "health.score.sum",adjusted="yes",sex="female")
```

## Chronic disease score
Men:  "chronic diseases" is associated with height loss, even after adjustment. Smoker and men with excess weight show a significantly higher risk for chronic diseases, than non-smoking men and non-excess weight men. Moreover men performing exercises more than once a week show a lower risk of getting chronic diseases
<br />
<br />
Women:  "chronic diseases" is associated with height loss, but not significant anymore after adjustment. Women eating fruits almost every day show a lower risk of chronic diseases compared to women eating no fruits. Women doing exercise show a lower risk of chronic diseases compared to women doing no exercise.

### Male
Unadjusted
```{r}
function_health_outcome(varOutcome = "chronicDiseaseScore",adjusted="no",sex="male")
```

adjusted
```{r}
function_health_outcome(varOutcome = "chronicDiseaseScore",adjusted="yes",sex="male")
```

### Female
Unadjusted
```{r}
function_health_outcome(varOutcome = "chronicDiseaseScore",adjusted="no",sex="female")
```

adjusted
```{r}
function_health_outcome(varOutcome = "chronicDiseaseScore",adjusted="yes",sex="female")
```

## Osteoarthritis
Men: "Osteoarthritis" is associated with height loss, even after adjustment.
Smoker and men with excess weight show a significantly higher risk for osteoarthritis, than non-smoking men and non-excess weight men. Moreover men performing exercises more than once a week show a lower risk of getting chronic diseases
<br />
<br />
Women: "Osteoarthritis" is associated with height loss, but not significant anymore after adjustment. Women doing exercise show a lower risk of Osteoarthritis compared to women doing no exercise.

### Male
Unadjusted
```{r}
function_health_outcome(varOutcome = "osteo",adjusted="no",sex="male")
```

adjusted
```{r}
function_health_outcome(varOutcome = "osteo",adjusted="yes",sex="male")
```

### Female
Unadjusted
```{r}
function_health_outcome(varOutcome = "osteo",adjusted="no",sex="female")
```

adjusted
```{r}
function_health_outcome(varOutcome = "osteo",adjusted="yes",sex="female")
```

## Ability to walk restricted due to pain
Men:  "Pain while walking" is associated with height loss, even after adjustment.
<br />
<br />
Women: "Pain while walking " is associated with height loss, even after adjustment.Women doing exercise show a lower risk of Osteoarthritis compared to women doing no exercise. Smoker and women with excess weight show a significantly higher risk for "Pain while walking"

### Male
Unadjusted
```{r}
function_health_outcome(varOutcome = "pain.walking",adjusted="no",sex="male")
```

adjusted
```{r}
function_health_outcome(varOutcome = "pain.walking",adjusted="yes",sex="male")
```

### Female
Unadjusted
```{r}
function_health_outcome(varOutcome = "pain.walking",adjusted="no",sex="female")
```

adjusted
```{r}
function_health_outcome(varOutcome = "pain.walking",adjusted="yes",sex="female")
```

